Adaptive Disk Storage for Interaction Graphs

### 2014-03-24 notes/graph-adapt/main.tex
added some notes

@@

\section{Introduction}
%
Typical interaction graphs have data associated with their vertices and edges.
Most algorithms access this data as they traverse the graph. However, not all
of the data is accessed by all the queries. Typically, there are correlations
among the attributes accessed by different queries, such as Q1 and Q5
accessing attributes a and b, and Q2, Q3, Q4 accessing attributes c and d, and
so on.
%

In an interaction graph database organized as blocks of temporal neighborlists
(as in my earlier paper), it is important to store the edge and vertex
properties locally. For  instance, if the edge properties are kept away in a
relational table, there will be almost no benefit to the locality
optimizations performed for block organization (discussed in my earlier
paper), as we would have to go back and forth between the disk blocks to
access the edge attributes. 
%

@@

>>>>

\usepackage{alltt}

@@

\section{Introduction}
An \emph{interaction graph} is an append-only graph, where new edges and
vertices are added as time progresses. 
%

%
Typical interaction graphs have data associated with their vertices and edges.
Most algorithms access this data as they traverse the graph. However, not all
of the data is accessed by all the queries. Typically, there are correlations
among the attributes accessed by different queries, such as Q1 and Q5
accessing attributes a and b, and Q2, Q3, Q4 accessing attributes c and d, and
so on.
%

In an interaction graph database organized as blocks of temporal neighborlists
(as in my earlier paper), it is important to store the edge and vertex
properties locally. For  instance, if the edge properties are kept away in a
relational table, there will be almost no benefit to the locality
optimizations performed for block organization (discussed in \cite{gedik14}), as we would have to go back and forth between the disk blocks to
access the edge attributes. 
%

@@

\begin{alltt}\scriptsize
- Work through a concrete example of a database and a 
   couple of queries that would change the layout on disk.

- An interaction graph is an append-only graph, where new 
   edges and vertices are added as time progresses.

- Typically, in graph dbs, if a vertex is visited during a 
   traversal, then it is likely that its neighbors will also 
    be visited

- So, we store we place highly connected neighbors in the 
    same disk block, to minimize disk I/O.

- Interaction graphs are temporal. The edges between vertices 
   change over time.

- For interaction graphs, we need to group edges spatially 
   and temporally.

- Vertices and edges have data associated with them.

- We want to store as much relevant data as possible in the 
   same block.

- We will store:
   - edges and vertices that are connected in the same time interval
   - edges and vertices that are highly connected (neighbors)
   - edge attributes that are likely to be accessed
   - vertex attributes that are likely to be accessed

- Example: We use an interaction graph to model connections
   between people on Twitter. Every time

- $C^d(B) = conductance, lower is better.$
- $C^h(B) = cohesiveness, higher is better.$

\end{alltt}

$L(B) = \sqrt[2]{C^h(B) \cdot ( 1 - C^d(B))}$
